---
title: "HCPA- Data Cleaning"
output:
  html_document:
    code_folding: hide
date: "2025-07-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
rm(list=ls())
library(R.matlab)
library(dplyr)
library(readxl)
library(knitr)
library(kableExtra)
library(arsenal)
library(stringr)
library(devtools)
devtools::install_github("strengejacke/strengejacke")
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(magick)
library(lme4)
library(lubridate)
library(tidyverse)
library(ggplot2)
```

## loading dataset
```{r}
segregation_index <- read.csv('~/Desktop/segregation_index_W_n725_2025_07_08.csv')
hcpa_merged = readxl::read_xlsx('~/Desktop/hcpa_merged_demo_cog_v1_2021_04_17.xlsx')
freesurfeer <- load('~/Desktop/HCPA_V1_n725_freesurfeer_DKTaseg_measures_2021_04_16.Rdat')
framwise <- read.csv('~/Desktop/Framewise_displacement_outlier_p3mm.csv')
ndar <- read.table("ndar_subject01.txt", header = TRUE, sep = "\t")
```

```{r}
library(tidyverse)
library(ggplot2)
library(ggpubr)

# Load your data (replace with your actual data loading method if needed)
# segregation_index <- read.csv("your_data.csv")

# Verify column names
print(colnames(segregation_index))

# Select columns for each atlas
power264 <- segregation_index %>% select(ID, contains("power264."))
sch200N7 <- segregation_index %>% select(ID, contains("scahefer200N7."))  # Note the spelling matches your data
sch200N17 <- segregation_index %>% select(ID, contains("scahefer200N17.")) # Note the spelling matches your data

# Define metrics
metrics <- c("W", "B", "S")

# Function to calculate correlations and create plots
analyze_atlas_pair <- function(atlas1, atlas2, name1, name2) {
  results <- list()
  
  for (metric in metrics) {
    # Construct column names
    col1 <- paste0(name1, ".", metric)
    col2 <- paste0(name2, ".", metric)
    
    # Extract data
    x <- atlas1[[col1]]
    y <- atlas2[[col2]]
    
    # Calculate correlation
    cor_test <- cor.test(x, y, method = "pearson")
    
    # Create scatter plot
    plot <- ggplot(data.frame(x, y), aes(x = x, y = y)) +
      geom_point(color = "blue", alpha = 0.6) +
      geom_smooth(method = "lm", color = "red", se = FALSE) +
      stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "top") +
      labs(title = paste(name1, "vs", name2, "-", metric),
           x = paste(name1, metric),
           y = paste(name2, metric)) +
      theme_minimal()
    
    # Store results
    results[[metric]] <- list(
      r = cor_test$estimate,
      p = cor_test$p.value,
      plot = plot
    )
  }
  
  return(results)
}

# Create list of atlases with correct names
atlases <- list(
  power264 = power264,
  scahefer200N7 = sch200N7,  # Matches your data spelling
  scahefer200N17 = sch200N17  # Matches your data spelling
)

# Generate all possible pairwise comparisons
atlas_pairs <- combn(names(atlases), 2, simplify = FALSE)

# Initialize results storage
correlation_results <- list()
all_plots <- list()

# Perform analyses
for (i in seq_along(atlas_pairs)) {
  pair <- atlas_pairs[[i]]
  name1 <- pair[1]
  name2 <- pair[2]
  
  cat("\nAnalyzing:", name1, "vs", name2, "\n")
  
  # Run analysis
  analysis <- analyze_atlas_pair(
    atlas1 = atlases[[name1]],
    atlas2 = atlases[[name2]],
    name1 = name1,
    name2 = name2
  )
  
  # Store results
  comparison_name <- paste(name1, "vs", name2)
  correlation_results[[comparison_name]] <- analysis
  
  # Collect plots
  for (metric in metrics) {
    plot_name <- paste(comparison_name, metric)
    all_plots[[plot_name]] <- analysis[[metric]]$plot
  }
}

# Print correlation results
cat("\nCorrelation Results:\n")
for (comparison in names(correlation_results)) {
  cat("\n", comparison, "\n")
  for (metric in metrics) {
    res <- correlation_results[[comparison]][[metric]]
    cat(metric, ": r =", round(res$r, 3), 
        ", p =", ifelse(res$p < 0.001, "<0.001", round(res$p, 3)), "\n")
  }
}

# Display all plots
cat("\nDisplaying plots...\n")
for (plot_name in names(all_plots)) {
  print(all_plots[[plot_name]])
}

```

By using Pearson correlation, all pairwise correlations for W, B, and S measures are exceptionally high (r > 0.94)
, and the p-values of 0 indicate these correlations are statistically significant beyond any reasonable doubt. The Schaefer atlas (200N7 or 200N17) seems to have the maximum consistency between Atlases, but Power264 is a more widely used in neuroimaging, even though it has a slightly lower (but still very strong) correlationos with Schaefer atlases. Thus, we choose Power264 atlas and W indices as our final decision.The W measure are not exactly same, but quite coherent.

## Create segregation dataset
```{r}
seg_data <- segregation_index %>%
  select(ID, power264.W, scahefer200N7.W, scahefer200N17.W)%>%
  rename(id = ID)
```

## Create framwise_clean dataset
```{r}
framwise_clean <- framwise %>%
  filter(perc.outliers <= 0.30)
```

## Create Grayratio
```{r}
fsdat$Grayratio <- fsdat$TotalGray/fsdat$eTIV

```

## Merge dataset
```{r}
hcpa_datav1 <- left_join(
  framwise_clean,                                 
  seg_data,         
  by = "id"                          
)
```

```{r}
hcpa_datav1 <- left_join(
  framwise_clean,                                 
  seg_data,         
  by = "id"                          
)
```

```{r}
hcpa_datav2 <- left_join(
  hcpa_datav1,                                 
  fsdat,         
  by = "id"                          
)
```

```{r}
hcpa_datav2 <- hcpa_datav2 %>%
  mutate(id = str_replace(id, "_V1_MR$", ""))
```

```{r}
hcpa_datav3 <- left_join(
  hcpa_datav2,                                 
  hcpa_merged,         
  by = c("id" = "src_subject_id")                          
)
```

```{r}
hcpa_datav4 <- left_join(
  hcpa_datav3,                                 
  ndar %>% select(src_subject_id, site),         
  by = c("id" = "src_subject_id")              
)
```


## Data Cleaning
```{r}
totalcomp_indices <- which(is.na(hcpa_datav4$nih_totalcogcomp_unadjusted))

# Print the indices
print(totalcomp_indices)
```

```{r}
fluid_indices <- which(is.na(hcpa_datav4$nih_fluidcogcomp_unadjusted))

# Print the indices
print(fluid_indices)
```

```{r}
cry_indices <- which(is.na(hcpa_datav4$nih_crycogcomp_unadjusted))

# Print the indices
print(cry_indices)
```

```{r}
hcpa_datav4$interview_age_years <- floor(hcpa_datav4$interview_age / 12)
age_indices <- which(is.na(hcpa_datav4$interview_age_years))

# Print the indices
print(age_indices)
```

```{r}
w_indices <- which(is.na(hcpa_datav4$power264.W))

# Print the indices
print(w_indices)
```

```{r}
ratio_indices <- which(is.na(hcpa_datav4$Grayratio))

# Print the indices
print(ratio_indices)
```

```{r}
sex_indices <- which(is.na(hcpa_datav4$sex))

# Print the indices
print(sex_indices)
```

```{r}
thickness_indices <- which(is.na(hcpa_datav4$MeanThickness))

# Print the indices
print(thickness_indices)
```

```{r}
edu_indices <- which(is.na(hcpa_datav4$moca_edu))

# Print the indices
print(edu_indices)
```

## Create final dataset
```{r}
hcpa_final <- hcpa_datav4 %>%
  filter(!is.na(nih_totalcogcomp_unadjusted) & !is.na(nih_fluidcogcomp_unadjusted) & !is.na(nih_crycogcomp_unadjusted) & !is.na(interview_age_years) & !is.na(power264.W) & !is.na(Grayratio) & !is.na(sex) & ! is.na(MeanThickness) &! is.na(interview_date)&! is.na(moca_edu))%>%
  mutate(interview_date = format(mdy(interview_date), "%Y-%m-%d"))
```

## Standardize W & Grayratio
```{r}
hcpa_final[, c("Grayratio_scale", "power264.W_scale", "scahefer200N7.W_scale", "scahefer200N17.W_scale")] <- 
  scale(hcpa_final[, c("Grayratio", "power264.W", "scahefer200N7.W", "scahefer200N17.W")])

```

```{r}
write.csv(hcpa_final, "~/Desktop/hcpa_final.csv", row.names = FALSE)
```


